---
title: ""
author: "Jong-Hoon Kim"
date: "1/23/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---
```{r, setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE, comment = '', fig.width = 6, fig.height = 6)
extrafont::loadfonts(device = "win", quiet = TRUE)# load fonts - every session
ggplot2::theme_set(theme(text = element_text(family = "Calibri", size = 12),
                         plot.title = element_text(size = 16, hjust = 0, vjust = 0.5),
                         plot.subtitle = element_text(size = 13),
                         plot.caption = element_text(size = 12)))
```


```{r data}
theta_ <- data.table::fread("inst/extdata/theta.csv")
theta <- theta_$val
names(theta) <- theta_$name
y0_ <- data.table::fread("inst/extdata/y0.csv")
y0 <- y0_$val
names(y0) <- y0_[["name"]]
 
usethis::use_data(y0, overwrite = T)
usethis::use_data(theta, overwrite = T)
 
devtools::document()
devtools::load_all(".")
```

# Check functions
## process_model
```{r}
np <- 100 # number of particles
tend <- 20
beta = theta[["R0"]] * theta[["gamma"]]
# volatility for beta (random walk)
beta_vol <- matrix(rnorm(np * tend, mean = 0, sd = theta[["betavol"]]), nrow = tend)
beta_vol[1,] <- exp(beta_vol[1,]) * beta # initial beta
y <- data.frame(matrix(0, nrow = np, ncol = length(y0)))
names(y) <- names(y0)
y[,"S"] <- 1e7
y[,"I"] <- 100
yt <- y1 <- process_model(params = theta, y = y, beta = beta_vol[1,])
for (i in 1:100) {
  yt <- process_model(params = theta, y = yt, beta = beta_vol[1,])
}
head(y1)
head(yt)
head(beta_vol[1,])
```

## Rt data generation
```{r}
nt <- 200 # days
param <- list()
param$sigma <- 1/5.2 # 1 / sigma = mean incubation period
param$epsilon <- 1/3 # 1/epsilon = mean latent period
param$gamma <- 1/4.3
param$R <- c(rep(1.2, nt/4), rep(1.4, nt/4), rep(0.8, nt/4), rep(1.2, nt/4))
param$presymp_infect <- FALSE
times <- 0:(nt-1) # simulation times
yini <- c(S = 1e7, E = 40, P = 40, I = 100, R = 0, CE = 0,  CI = 0, CR = 0) # initial values

library(deSolve) 
library(tidyverse)

ode(func = sepir, y = yini, times = times, parms = param) %>%
  as.data.frame() -> out

inf_daily <- diff(out$CE)
onset_daily <- diff(out$CI)
confirm_daily <- diff(out$CR)

df <- data.frame(t = 1:length(param$R), 
                 daily_R_true = param$R, 
                 daily_infect = c(0, inf_daily),
                 daily_onset = c(0, onset_daily),
                 daily_confirm = c(0, confirm_daily))

df %>% 
  pivot_longer(cols = -t) %>%
  filter(name != "daily_R_true") %>% 
  ggplot(aes(t, value, color = name)) +
  geom_line() +
  labs(x = "time (day)", y = "number of individuals", color = "")

# Rt_data <- df
# usethis::use_data(Rt_data, overwrite = T)
```

## particle_filter - sample
```{r}
pf <- particle_filter(data_type = "infection", npart = 1e4, dt = 0.01, error_pdf = "pois")
pf$lik_overall_average
df <- cbind(Rt_data[,1:2], Rt_pf = pf$trace$beta/theta[["gamma"]])
suppressMessages(library(tidyverse))
df %>% pivot_longer(cols = -t) %>% 
  ggplot(aes(t, value, color = name)) + 
  geom_line()
```

## particle_filter - filtered distribution
```{r}
theta[["betavol"]] = 0.4
pf <- particle_filter(data_type = "infection", npart = 1e4, dt = 0.02, error_pdf = "negbin", negbin_size = 100)
pf$lik_overall_average
pr <- c(0.025, 0.25, 0.5, 0.75, 0.975)

Rt_quantile <- as.data.frame(t(apply(pf$beta_filtered/theta[["gamma"]], 1, function(x) {quantile(x, pr)})))
Rt_mean <- data.frame(mean = rowMeans(pf$beta_filtered/theta[["gamma"]]))
df <- cbind(Rt_quantile, Rt_mean, Rt_data[,1:2])

suppressMessages(library(tidyverse))
ggplot(df, aes(x = t)) +
  geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), fill = "grey") +
  geom_line(aes(y = `50%`), size = 1) + 
  geom_line(aes(y = mean), color = "steelblue", size = 1) + 
  geom_point(aes(y = daily_R_true), color = "darkred", size = 1) +
  labs(title = "R(t) via particle filtering", y = "R(t)", x = "Day")

plot_latent_var(sim_res = pf, data_type = "infection")
```

## Particle filtering and backward sampling
\[X(t_k) | Y(t_1)=y^*_1, \dots,\; Y(t_n)=y^*_n,\; X_k | Y_1=y_1^*, \dots,\; Y_n=y_n^*\]

## repeat particle_filter and do backward sampling 
```{r}
res <- run_model(data_type = "infection", rep = 1e2, npart = 1e3, dt = 0.1,
                 systematic_resampling = F)
          
pr <- c(0.025, 0.25, 0.5, 0.75, 0.975)
Rt_quantile <- as.data.frame(t(apply(res$Rt, 1, function(x) {quantile(x, pr)})))
library(data.table)
Rt_mean <- data.frame(mean = rowMeans(res$Rt), 
                      rollmean = frollmean(Rt_quantile[,3], n=7))
df <- cbind(Rt_quantile, Rt_mean, Rt_data[,1:2])
suppressPackageStartupMessages(library(tidyverse))
ggplot(df, aes(x = t)) +
  geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), fill = "grey") +
  geom_line(aes(y = `50%`), size = 1) + 
  geom_line(aes(y = mean), color = "steelblue", size = 1) + 
  geom_point(aes(y = daily_R_true), color = "darkred") +
  labs(title = "R(t) via particle filtering with backward sampling", y = "R(t)", x = "Day")
```

## Parallel forward particle filtering and backward sampling 
```{r}
library(parallel)
library(doParallel)
library(foreach)
ncores = detectCores()
cl <- makeCluster(getOption("cl.cores", ncores-2))
doParallel::registerDoParallel(cl)

system.time({
r <- foreach (i = 1:1e3, .combine = cbind, .packages ="pfilter", .inorder = F) %dopar% {
  run_model(data_type = "infection", rep = 1, npart = 1e4, tend = 200, dt = 0.02, systematic_resampling = FALSE)
}})

parallel::stopCluster(cl)

rt = r[9,] # Rt
rtdf = data.frame(matrix(unlist(rt), ncol=length(rt), byrow=F))
pr <- c(0.025, 0.25, 0.5, 0.75, 0.975)
Rt_quantile <- as.data.frame(t(apply(rtdf, 1, function(x) {quantile(x, pr)})))
library(data.table)
Rt_mean <- data.frame(mean = rowMeans(rtdf), 
                      rollmean = frollmean(Rt_quantile[,3], n=7))
df <- cbind(Rt_quantile, Rt_mean, Rt_data[,1:2])
suppressPackageStartupMessages(library(tidyverse))
ggplot(df, aes(x = t)) +
  geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), fill = "grey") +
  geom_line(aes(y = `50%`), size = 1) + 
  geom_line(aes(y = mean), color = "steelblue", size = 1) + 
  geom_point(aes(y = daily_R_true), color = "darkred", size = 1) +
  labs(title = "R(t) via particle filtering with backward sampling", y = "R(t)", x = "Day")

plot_latent_var_pl(sim_res = r, data_type = "infection")
```


## parallel repeat particle_filter and do backward sampling 
```{r}
devtools::load_all(".")
library(parallel)
library(doParallel)
library(foreach)
ncores = detectCores()
cl <- makeCluster(getOption("cl.cores", ncores-2))
doParallel::registerDoParallel(cl)
system.time({
r <- foreach (i=1:1e3, .combine = cbind, .packages = "pfilter", .inorder = F) %dopar% {
  run_model(params = theta, y = y0, data = Rt_data, data_type = "confirmation",
            rep = 1, npart = 1e3, tend = 200, dt = 0.1, 
            systematic_resampling = FALSE)
}})
parallel::stopCluster(cl)

rt <- r[9,] # Rt
rtdf = data.frame(matrix(unlist(rt), ncol=length(rt), byrow=F))
pr <- c(0.025, 0.25, 0.5, 0.75, 0.975)
Rt_quantile <- as.data.frame(t(apply(rtdf, 1, function(x) {quantile(x, pr)})))
library(data.table)
Rt_mean <- data.frame(mean = rowMeans(rtdf), 
                      rollmean = frollmean(Rt_quantile[,3], n=7))
df <- cbind(Rt_quantile, Rt_mean, Rt_data[,1:2])
suppressPackageStartupMessages(library(tidyverse))
ggplot(df, aes(x = t)) +
  geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), fill = "grey") +
  geom_ribbon(aes(ymin = `25%`, ymax = `75%`), fill = "grey50") +
  geom_line(aes(y = `50%`), size = 1) + 
  geom_line(aes(y = mean), color = "steelblue", size = 1) + 
  geom_line(aes(y = daily_R_true), color = "darkred") +
  labs(title = "R(t) via particle filtering with backward sampling", y = "R(t)", x = "Day")

plot_latent_var_pl(sim_res = r, data_type = "confirmation")
```

## Maximum likelihood estimation test
```{r}
# devtools::load_all(".")
mle <- maxlik(data_type = "confirmation")

df <- cbind(Rt_data[,1:2], (mle$beta_trace/theta[name == "gamma", val]))
suppressPackageStartupMessages(library(tidyverse))
df %>% 
  pivot_longer(cols = - t) %>%
  ggplot(aes(t, value, color = name)) +
  geom_line() +
  geom_point() +
  labs(title = "R(t) via maximum likelihood", y = "R(t)", x = "Day")

```

## run_step function 
```{r eval=F}
# nc <- 1
# S <- rep(1e3, nc)
# E <- rep(10, nc)
# P <- rep(10, nc)
# I <- rep(10, nc)
# R <- rep(0, nc)
# daily_infect <- rep(0, nrow(y))
# daily_symptom <- rep(0, nrow(y))
# daily_confirm <- rep(0, nrow(y))
# 
# yini <- data.frame(S = S, E = E, P = P, I = I, R = R, CE = daily_infect, CI = daily_symptom, CR = daily_confirm)
# 
# run_step <- function(y, times, params){
#   param$presymp_infect <- FALSE
#   out <- as.data.frame(ode(y = y, func = sepir, times = times, parms = param))
#   tail(out, 1)
# }
#     
# # yini <- c(S = 1e7, E = 40, P = 40, I = 100, R = 0, CE = 0,  CI = 0, CR = 0) # initial values
# out <- apply(yini, 1, run_step, times = c(0, 1), params = as.list(theta))
# outdf <- data.frame(matrix(unlist(out), nrow = length(out), byrow = T))
# outdf <- outdf[, -1] # remove the time column
# names(outdf) <- names(yini)
```


## KDCA data
### Data cleaning
```{r}
library(readxl)
suppressMessages(library(tidyverse))
d <- read_xlsx("covid_kdca.xlsx")
names(d) = c("id", "nationality", "sex", "age", "sido", "sigungu", "date_symptom_onset", "date_diagnosis", "date_report", "route_infection", "country_orgin", "id_infector", "occupation")
d1 <- d %>% filter(route_infection == "국내")%>% 
  select(date_diagnosis) %>% 
  group_by(date = date_diagnosis) %>% 
  summarise(daily_confirm = n())
d2 <- data.frame(date = seq(as.Date("2020-01-03"), as.Date("2021-02-23"), by = "day"))
d2$t <- 1:nrow(d2)
d3 <- left_join(d2, d1, by = "date")
kdca_dat <- d3 %>% mutate(daily_confirm = ifelse(is.na(daily_confirm), 0, daily_confirm))

# saveRDS(dat, paste0( "outputs/dat_", tstamp, ".rds"))
## name standardization
ggplot(kdca_dat, aes(t, daily_confirm)) +
  geom_line()

# library(data.table)
# dat$daily_confirm <- frollmean(dat$daily_confirm, n = 7)
# dat <- dat[7:nrow(dat),]
# dat$t <- 1:nrow(dat)
# dat$daily_confirm <- round(dat$daily_confirm)
```



### Data fitting - forward filtering
```{r}
dat <- kdca_dat
theta[["betavol"]] = 0.55
ynew <- c(S = 5*1e7, E = 1e1, P = 1e1, I = 1e1, R = 0, CE = 0, CI = 0, CR = 0)
set.seed(10)
pf <- particle_filter(y = ynew, data = dat, data_type = "confirmation", 
                      npart = 2e4, tend = nrow(dat), dt = 0.2, 
                      error_pdf = "negbin", negbin_size = 20)

pf$lik_overall_average
pr <- c(0.025, 0.25, 0.5, 0.75, 0.975)
Rt_quantile <- as.data.frame(t(apply(pf$beta_filtered/theta[["gamma"]], 1, 
                                     function(x) {quantile(x, pr)})))
Rt_mean <- data.frame(mean = rowMeans(pf$beta_filtered/theta[["gamma"]]))
df <- cbind(Rt_quantile, Rt_mean, dat[,c("t", "date")])
df$date <- seq(as.Date("2020-01-20"), length.out = nrow(df), by = "day")
suppressMessages(library(tidyverse))

ggplot(df, aes(x = date)) +
  geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), fill = "grey") +
  geom_ribbon(aes(ymin = `25%`, ymax = `75%`), fill = "grey60") +
  geom_line(aes(y = `50%`), color = "steelblue", size = 1, linetype = "dotted") + 
  geom_line(aes(y = mean), color = "steelblue", size = 1) + 
  labs(title = "R(t) via particle filtering", y = "R(t)", x = "") +
  geom_hline(yintercept = 1, color = "black", linetype = "dotted") + 
  scale_y_continuous(limits = c(0, 10), breaks = c(0, 1, seq(2, 10, 2)), 
                     labels = c(0, 1, seq(2, 10, 2))) + 
  scale_x_date(date_breaks = "2 month", date_minor_breaks = "1 month", 
               limits = c(as.Date("2020-01-20"), max(df$date)))

plot_latent_var(sim_res = pf, data = dat, data_type = "confirmation")
```

### Parallel - forward filtering with backward sampling
```{r}
y0 <- c(S = 5*1e7, E = 1e1, P = 1e1, I = 1e1, R = 0, CE = 0, CI = 0, CR = 0)
theta["betavol"] <- 0.6
usethis::use_data(y0, overwrite = T)
usethis::use_data(theta, overwrite = T)
devtools::load_all(".")

library(parallel)
library(doParallel)
library(foreach)
ncores = detectCores()
cl <- makeCluster(getOption("cl.cores", ncores-2))
doParallel::registerDoParallel(cl)

system.time({
r <- foreach (i = 1:1e3, .combine = cbind, .packages ="pfilter", .inorder = F) %dopar% {
  run_model(params = theta, y = y0, data = dat, data_type = "confirmation",
            rep = 1, npart = 1e4, tend = nrow(dat), dt = 0.1, error_pdf = "negbin", negbin_size = 10,
            systematic_resampling = FALSE)
}})
parallel::stopCluster(cl)

# tstamp <- format(Sys.time(), "%Y%m%dT%H%M%S")
# saveRDS(r, paste0( "outputs/ffbs_", tstamp, ".rds"))
tstamp <- "20210305T141259"
r <- readRDS(paste0( "outputs/ffbs_", tstamp, ".rds"))
dat <- readRDS(paste0( "outputs/dat_", tstamp, ".rds"))


rt <- r[9,] # Rt
rtdf = data.frame(matrix(unlist(rt), ncol=length(rt), byrow=F))
pr <- c(0.025, 0.25, 0.5, 0.75, 0.975)
Rt_quantile <- as.data.frame(t(apply(rtdf, 1, function(x) {quantile(x, pr)})))
library(data.table)
Rt_mean <- data.frame(mean = rowMeans(rtdf), 
                      rollmean = frollmean(Rt_quantile[,3], n=7))
df <- cbind(Rt_quantile, Rt_mean, dat[, 2:3])
df$date <- seq(as.Date("2020-01-20"), length.out = nrow(df), by = "day")
suppressMessages(library(tidyverse))

ggplot(df, aes(x = date)) +
  geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), fill = "steelblue", alpha = 0.3) +
  geom_ribbon(aes(ymin = `25%`, ymax = `75%`), fill = "steelblue", alpha = 0.6) +
  geom_line(aes(y = `50%`), color = "steelblue", size = 1, linetype = "dotted") + 
  geom_line(aes(y = mean), color = "steelblue", size = 1) + 
  labs(title = "R(t) via particle filtering with backward sampling", y = "R(t)", x = "Day") +
  geom_hline(yintercept = 1, color = "darkred", size = 1, linetype = "dotted") +
  scale_y_continuous(limits = c(0, 10), breaks = c(0, 1, seq(2, 10, 2)), labels = c(0, 1, seq(2, 10, 2))) + 
  scale_x_date(date_breaks = "2 month", date_minor_breaks = "1 month", limits = c(as.Date("2020-01-20"), max(df$date))) +
  labs(x = "", y = "R(t)") 
  # theme(axis.text.x = element_text(angle = 0, vjust = 1.0, hjust = 0.5))
# ggsave("plots/Rt_ffbs.png", width = 16, height = 9, scale = 1, units = "cm")
# plt <- plot_latent_var_pl(sim_res = r, data = dat, data_type = "confirmation")
# ggsave("plots/daily_confirm_ffbs.png", plt, width = 16, height = 9, scale = 1, units = "cm")
```

## Can we stop and continue at any arbitrary time step?
That is, can we fun the PF 
1. particle_filter function that takes the weight (W) and particle ID (A) matrix, 
beta, and latent variable

```{r}
theta[["betavol"]] = 0.5
ynew <- c(S = 5*1e7, E = 1e1, P = 1e1, I = 1e1, R = 0, CE = 0, CI = 0, CR = 0)
set.seed(10)
tbegin <- 100
tend <- 100
dat <- kdca_dat[tbegin:(tend+tbegin-1),]
y_pf <- as.data.frame(pf$latent_var_filtered[,tend,])
names(y_pf) <- names(y0)
beta_pf <- pf$beta_filtered[tend,]

pf <- particle_filter(y = y_pf, beta0 = beta_pf,
                      data = dat, data_type = "confirmation", 
                      npart = 2e4, tend = tend, dt = 0.1,
                      error_pdf = "negbin", negbin_size = 15)
 
pf$lik_overall_average
pr <- c(0.025, 0.25, 0.5, 0.75, 0.975)
Rt_quantile <- as.data.frame(t(apply(pf$beta_filtered/theta[["gamma"]], 1, 
                                     function(x) {quantile(x, pr)})))
Rt_mean <- data.frame(mean = rowMeans(pf$beta_filtered/theta[["gamma"]]))
df <- cbind(Rt_quantile, Rt_mean, dat[,c("t", "date")])
# df$date <- seq(as.Date("2020-01-03") + tbegin - 1, length.out = nrow(df), by = "day")

suppressMessages(library(tidyverse))

ggplot(df, aes(x = date)) +
  geom_ribbon(aes(ymin = `2.5%`, ymax = `97.5%`), fill = "steelblue", alpha = 0.3) +
  geom_ribbon(aes(ymin = `25%`, ymax = `75%`), fill = "steelblue", alpha = 0.6) +
  geom_line(aes(y = `50%`), color = "steelblue", size = 1, linetype = "dotted") + 
  geom_line(aes(y = mean), color = "steelblue", size = 1) + 
  labs(title = "R(t) via particle filtering", y = "R(t)", x = "") +
  geom_hline(yintercept = 1, color = "red", linetype = "dotted") + 
  # scale_y_continuous(limits = c(0, 10), breaks = c(0, 1, seq(2, 10, 2)), 
  #                    labels = c(0, 1, seq(2, 10, 2))) + 
  scale_x_date(date_breaks = "2 month", date_minor_breaks = "1 month", 
               limits = c(min(df$date), max(df$date)))

plot_latent_var(sim_res = pf, data = dat, data_type = "confirmation")

```

## Testing arma::cube can be access through _ just like in Rcpp::NumericMatrix
```{r}
library(Rcpp)
library(RcppArmadillo)
sourceCpp("cpp/test.cpp")
# a = array(1:12, c(2,3,2))
# a[, , 1]
# (a = export_array(Q = a, slice = 0))
# # (m = export_mat(Q = a, slice = 0))
# (v = export_vec(Q = a, col = 0, slice = 0))
# (i = to_int(x = c(1.0, 2.2)))
# m = matrix(1:12, nrow = 3)
# m[,1]
# (v = export_vec_from_mat(m = m, col = 0))
# (ch = add_x("mr. "))
# all_x("x")
# all_x("y")
int_sample(10)
int_wt_sample(10, wt = c(0.8,0.2,rep(0,8)))
sample_test(10, wt = c(0.8,0.2,rep(0,8)))
```

## testing assign_weights function
```{r}
library(Rcpp)
library(RcppArmadillo)
sourceCpp("cpp/assign_weights_cpp.cpp")
st_now = data.matrix(data.frame(S=rep(1,10), E1 = rep(1,10), E2 = rep(1,10), I = rep(1,10), R = 1:10, CE1 = 101:110, CI = 101:110))
st_before = data.matrix(data.frame(S=rep(1,10), E1 = rep(1,10), E2=rep(1,10), I = rep(1,10), R = 1:10-1, CE1 = round(seq(1, 100, length.out = 10)), CI = round(seq(1, 100, length.out = 10))))

wt = assign_weights_cpp(st_now, st_before, 53L, data.matrix(Rt_data), "infection")
all.equal(wt[1], dpois(101,100))
```

## testing particle_filter_cpp function
```{r}
library(Rcpp)
library(RcppArmadillo)
sourceCpp("cpp/particle_filter_cpp.cpp")
```

##Conditional log likelihood
The estimated conditional log likelihood from a fitted model.
The conditional likelihood is defined to be the value of the density
\[ Y(t_k) | Y(t_1),\dots,Y(t_{k-1}) Y_k | Y_1,\dots,Y_{k-1}\] 
evaluated at \[Y(t_k) = y^*_k Y_k = y_k^*\].

Here, \({Y(t_k)}{Yk}\) is the observable process, and \(y^*_kyk*}\) the data, at time \(t_k\).

Thus the conditional log likelihood at time \(t_k\) is
\[{\ell_k(\theta) = \log f[Y(t_k)=y^*_k \vert Y(t_1)=y^*_1, \dots, Y(t_{k-1})=y^*_{k-1}],}{ell_k(theta)=log f[Yk = yk* | Y1=y1*, \dots, Y(k-1)=y(k-1)*],}\]
where is the probability density above.

## pomp package
```{r}
library(tidyverse)
read_csv(paste0("https://kingaa.github.io/sbied/stochsim/", "Measles_Consett_1948.csv")) %>%
select(week,reports=cases) -> meas
meas %>% as.data.frame() %>% head()

sir_step <- function (S, I, R, N, Beta, mu_IR, delta.t, ...) {
  dN_SI <- rbinom(n=1,size=S,prob=1-exp(-Beta*I/N*delta.t))
  dN_IR <- rbinom(n=1,size=I,prob=1-exp(-mu_IR*delta.t))
  S <- S - dN_SI
  I <- I + dN_SI - dN_IR
  R <- R + dN_IR
  c(S = S, I = I, R = R)
}

sir_rinit <- function (N, eta, ...) {
  c(S = round(N*eta), I = 1, R = round(N*(1-eta)))
}

library(pomp)
meas %>%
pomp(times="week",t0=0,
rprocess=euler(sir_step,delta.t=1/7),
rinit=sir_rinit
) -> measSIR

sir_dmeas <- function (reports, H, rho, log, ...) {
  dbinom(x=reports, size=H, prob=rho, log=log)
}
sir_rmeas <- function (H, rho, ...) {
  c(reports=rbinom(n=1, size=H, prob=rho))
}
measSIR %>%
  pomp(
  rmeasure=sir_rmeas,
  dmeasure=sir_dmeas
) -> measSIR

measSIR %>%
  simulate(
  params=c(Beta=7.5,mu_IR=0.5,rho=0.5,eta=0.03,N=38000),
  nsim=20,format="data.frame",include.data=TRUE
) -> sims
sims %>%
  ggplot(aes(x=week,y=reports,group=.id,color=.id=="data"))+
  geom_line()+
  guides(color=FALSE)
```

```{r}
library(pomp)
library(tidyverse)
seir_step <- Csnippet("
  double N = S + E1 + E2 + I + R;
  double dN_SE1 = beta * S * I / N * dt;
  double dN_E1E2 = 2 * sigma * E1 * dt;
  double dN_E2I = 2 * sigma * E2 * dt;
  double dN_IR = gamma * I * dt;
  S +=  - dN_SE1;
  E1 += dN_SE1 - dN_E1E2;
  E2 += dN_E1E2 - dN_E2I;
  I += dN_E2I - dN_IR;
  R += dN_IR;
  H += dN_SE1;
")

seir_rinit <- Csnippet("
  S = S_0;
  E1 = E1_0;
  E2 = E2_0;
  I = I_0;
  R = R_0;
  H = H_0;
")

seir_dmeas <- Csnippet("
  lik = dpois(daily_infect, H, give_log);
")

seir_rmeas <- Csnippet("
  daily_infect = rpois(H);
")

dat <- data.frame(day = Rt_data$t, daily_infect = Rt_data$daily_infect)

dat %>% pomp(
  times = "day", 
  t0 = 1,
  rprocess = euler(seir_step, delta.t = 0.2),
  rinit = seir_rinit,
  rmeasure = seir_rmeas,
  dmeasure = seir_dmeas,
  accumvars = "H",
  statenames = c("S", "E1", "E2", "I", "R", "H"),
  paramnames = c("beta", "sigma", "gamma", "S_0", "E1_0", "E2_0", "I_0", "R_0", "H_0")) -> covid_seir

covid_seir %>%
  simulate(params = c(beta=1.2*0.25, sigma=0.25, gamma=0.25,
                      S_0=1e7, E1_0=40, E2_0=40, I_0=100, R_0=0, H_0=0),
nsim = 20, format="data.frame", include.data=TRUE) -> sims

sims %>%
  ggplot(aes(x=day,y=daily_infect, group=.id, color=.id=="data"))+
  geom_line()+
  guides(color=FALSE)

covid_seir %>%
  pfilter(Np=1000, paramnames="beta") -> pfrick

```

```{r}
library(pomp)
library(tidyverse)
seir_step <- Csnippet("
  double N = S + E1 + E2 + I + R;
  double dN_SE1 = beta * S * I / N * dt;
  double dN_E1E2 = 2 * 0.25 * E1 * dt;
  double dN_E2I = 2 * 0.25 * E2 * dt;
  double dN_IR = 0.25 * I * dt;
  S +=  - dN_SE1;
  E1 += dN_SE1 - dN_E1E2;
  E2 += dN_E1E2 - dN_E2I;
  I += dN_E2I - dN_IR;
  R += dN_IR;
  H += dN_SE1;
")

seir_rinit <- Csnippet("
  S = 1e7;
  E1 = 40;
  E2 = 40;
  I = 100;
  R = 0;
  H = 0;
")

seir_dmeas <- Csnippet("
  lik = dpois(daily_infect, H, give_log);
")

seir_rmeas <- Csnippet("
  daily_infect = rpois(H);
")

dat <- data.frame(day = Rt_data$t, daily_infect = as.integer(Rt_data$daily_infect))

dat %>% pomp(
  times = "day", 
  t0 = 1,
  rprocess = euler(seir_step, delta.t = 0.2),
  rinit = seir_rinit,
  rmeasure = seir_rmeas,
  dmeasure = seir_dmeas,
  accumvars = "H",
  statenames = c("S", "E1", "E2", "I", "R", "H"),
  paramnames = c("beta")) -> covid_seir

covid_seir %>%
  simulate(params = c(beta=1.2*0.25),
nsim = 20, format="data.frame", include.data=TRUE) -> sims

sims %>%
  ggplot(aes(x=day,y=daily_infect, group=.id, color=.id=="data"))+
  geom_line()+
  guides(color=FALSE)

pf = pfilter(covid_seir, Np=1000, params = c(beta = 0.3), 
             filter.mean = T, filter.traj = T, verbose = T) 
pf@filter.mean %>% 
  t() %>%
  as.data.frame() %>%
  mutate(time = 1:n()) %>% 
  pivot_longer(-time) %>%
  filter(name == "I") %>% 
  ggplot(aes(time, value)) +
  geom_line(aes(color = name))
  
```      

```{r}
ewmeas %>%
  subset(time < 1952) %>%
  pomp(
    times="time",t0=1948,
    rprocess=euler(
      Csnippet("
        int nrate = 6;
        double rate[nrate];	// transition rates
        double trans[nrate];	// transition numbers
        double dW;

        // gamma noise, mean=dt, variance=(sigma^2 dt)
        dW = rgammawn(sigma,dt);

        // compute the transition rates
        rate[0] = mu*pop;	// birth into susceptible class
        rate[1] = (iota+Beta*I*dW/dt)/pop; // force of infection
        rate[2] = mu;		// death from susceptible class
        rate[3] = gamma;	// recovery
        rate[4] = mu;		// death from infectious class
        rate[5] = mu; 		// death from recovered class

        // compute the transition numbers
        trans[0] = rpois(rate[0]*dt);	// births are Poisson
        reulermultinom(2,S,&rate[1],dt,&trans[1]);
        reulermultinom(2,I,&rate[3],dt,&trans[3]);
        reulermultinom(1,R,&rate[5],dt,&trans[5]);

        // balance the equations
        S += trans[0]-trans[1]-trans[2];
        I += trans[1]-trans[3]-trans[4];
        R += trans[3]-trans[5];
      "),
      delta.t=1/52/20
    ),
    rinit=Csnippet("
      double m = pop/(S_0+I_0+R_0);
      S = nearbyint(m*S_0);
      I = nearbyint(m*I_0);
      R = nearbyint(m*R_0);
    "),
    paramnames=c("mu","pop","iota","gamma","Beta","sigma",
      "S_0","I_0","R_0"),
    statenames=c("S","I","R"),
    params=c(mu=1/50,iota=10,pop=50e6,gamma=26,Beta=400,sigma=0.1,
      S_0=0.07,I_0=0.001,R_0=0.93)
  ) -> ew1
```

