---
title: "Reproduction number"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

## Install EpiEstim package if it is not already
```{r}
list_of_packages <- c("EpiEstim") # packages we need
new_packages <- list_of_packages[!(list_of_packages %in% installed.packages()[,"Package"])] 
if(length(new_packages)) install.packages(new_packages)
```

## Global parameters
```{r}
# generation time follows a Gamma distribution
gt_shape <- 2 # shape parameter
gt_rate <- 1/2 # rate parameter
# incubation period (infection - onset of symptoms) follows a Gamma distribution.
incu_shape <- 2 # shape parameter
incu_rate <- 1/3 # rate parameter
# delay from infection to confirmation follows a Gamma distribution.
conf_shape <- 2 # shape parameter
conf_rate <- 1/4 # rate parameter
# Rt time series
Rt <- c(rep(1.2, 40), rep(2.2, 6), rep(0.8, 6), rep(0.8, 20))
```

## Create fake data using the renewal equation
```{r}
## weight function = generation time distribution
wt <- function(x){
  dgamma(x, shape = gt_shape, rate = gt_rate)
}
## weight function = generation time distribution
## this function give the number of people effective at timt t by accounting
## for all past infections and generation time distribution
infect_wt <- function(t, It) {
  inf_wt <- 0
  for(i in 1:(t-1)) {
    inf_wt <- inf_wt + It[t-i] * wt(i)
  }
  return(inf_wt)
}
# create_It
# This function creates infection time series where I0 is infecteds on day 1
create_It <- function(t = length(Rt)+1, I0 = 10) {
  It <- c(I0, rep(NA, t-1))
  for (i in 2:length(It)) {
    It[i] <- Rt[i-1] * infect_wt(i, It)
  }
  return(round(It))
}
# create infection time series and plot
It <- create_It()
plot(It, type = "l")
plot(Rt, type = "l")
```

## Rt estimation using EpiEstim package
```{r}
library(EpiEstim)
## infection time series. Column name is I
dat <- data.frame(I = It)

## this is to define a window for running average -- default is 7
window <- 7
tstart <- seq(3, nrow(dat) - window)
tend <- tstart + window

cori_R <- 
  estimate_R(incid = dat, method = c("parametric_si"),
             config = make_config(list(mean_si = gt_shape / gt_rate,
                             std_si = sqrt(gt_shape / (gt_rate^2)), 
                             t_start = tstart, t_end = tend)))
plot(1:length(Rt), Rt, col = 2)
lines(tstart, cori_R$R$`Median(R)`, type = "l")
```

## Compare with the [WHO-PAHO estimator](https://harvardanalytics.shinyapps.io/covid19/?utm_source=pocket_mylist)
This website does not seem to be working.
```{r}
d <- data.frame(date = seq(as.Date("2021-01-01"),  
                           by = "day", length.out = length(It)), 
                I = It)
dt <- transform(d, date = format(date, "%m/%d/%Y"))
write.csv(dt, file = "iseries.csv", row.names = F)
```

## Create onset time series
```{r}
# create_onset
# create time series of symptom onset based on the infection time series and 
# the pre-defined incubation period distribution
# if stoch = TRUE, incubation period is a Gamma random variable with shape = 
# incu_shape and rate = incu_rate
create_onset <- function(inf = NULL, stoch = FALSE){
  onset <- data.frame(t = rep(NA, sum(inf$inf)))
  id <- 1
  for (i in 1:nrow(inf)){
    for(j in 1:inf$inf[i]){
      if(stoch){
        onset$t[id] <- inf$t[i] + rgamma(1, incu_shape, incu_rate)
      } else {
        onset$t[id] <- inf$t[i] + incu_shape / incu_rate
      }
      id <- id + 1
    }
  }
  onset <- as.data.frame(table(round(onset)))
  names(onset) <- c("t", "freq")
  onset$t <- as.integer(as.character(onset$t))
  return(onset)
}

set.seed(3) # to make stochastic clinical onset time series reproducible 
inf <- data.frame(t = 1:length(It), inf = It)
onset <- create_onset(inf = inf, stoch = F)

plot(inf$t, inf$inf, xlim = c(0, 90), type = "l")
lines(onset$t, onset$freq, col = 2)
```

## Rt estimation on serial interval using EpiEstim package
```{r}
library(EpiEstim)
dat1 <- data.frame(t = 1:max(onset$t), I = 0)
dat2 <- data.frame(t = onset$t, I = onset$freq)

for (i in 1:nrow(dat1)) {
  id <- which(dat2$t == i)
  if(length(id) > 0){
    dat1[i, 2] <- dat2[id, 2]
  }
}

dat <- data.frame(I = round(dat1$I))

window <- 7
tstart <- seq(3, nrow(dat)- window )
tend <- tstart + window
cori_R <- 
  estimate_R(incid = dat, method=c("parametric_si"), 
             config = make_config(list(mean_si = gt_shape / gt_rate, 
                                       std_si = sqrt(gt_shape / (gt_rate^2)), 
                                       t_start = tstart, t_end= tend)))
# plot(cori_R)
plot(1:length(Rt), Rt, col = 2)
lines(tstart, cori_R$R$`Median(R)`, type = "l")
lines(tstart, cori_R$R$`Quantile.0.025(R)`, type = "l")
lines(tstart, cori_R$R$`Quantile.0.975(R)`, type = "l")
```

## Create time series of patient confirmation
```{r}
# create_confirm
# same as the create_onset with different parameters for Gamma distribution
# create time series of confirmation based on the infection time series and 
# the pre-defined delay distribution
# if stoch = TRUE, delay distribution is a Gamma random variable with shape = 
# conf_shape and rate = conf_rate
create_confirm <- function(inf = NULL, stoch = FALSE){
  confirm <- data.frame(t = rep(NA, sum(inf$inf)))
  id <- 1
  for (i in 1:nrow(inf)){
    for(j in 1:inf$inf[i]){
      if(stoch){
        confirm$t[id] <- inf$t[i] + rgamma(1, conf_shape, conf_rate)
      } else {
        confirm$t[id] <- inf$t[i] + conf_shape / conf_rate
      }
      id <- id + 1
    }
  }
  
  confirm <- as.data.frame(table(round(confirm)))
  names(confirm) <- c("t", "freq")
  confirm$t <- as.integer(as.character(confirm$t))
  return(confirm)
}
set.seed(3) # to make stochastic time series reproducible 
inf <- data.frame(t = 1:length(It), inf = It)
conf <- create_confirm(inf = inf)

plot(inf$t, inf$inf, xlim = c(0, 90), type = "l")
lines(conf$t, conf$freq, col = 2)
```

```{r}
library(EpiEstim)
dat1 <- data.frame(t = 1:max(conf$t), I = 0)
dat2 <- data.frame(t = conf$t, I = conf$freq)
for (i in 1:nrow(dat1)) {
  id <- which(dat2$t == i)
  if(length(id) > 0){
    dat1[i, 2] <- dat2[id, 2]
  }
}

dat <- data.frame(I = round(dat1$I))

window <- 1
tstart <- seq(3, nrow(dat)- window )
tend <- tstart + window
cori_R <- 
  estimate_R(incid = dat, method=c("parametric_si"), 
             config = make_config(list(mean_si = gt_shape / gt_rate, 
                                       std_si = sqrt(gt_shape / (gt_rate^2)), 
                                       t_start = tstart, t_end= tend)))

plot(1:length(Rt), Rt, col = 2)
lines(tstart, cori_R$R$`Median(R)`, type = "l")
lines(tstart, cori_R$R$`Quantile.0.025(R)`, type = "l")
lines(tstart, cori_R$R$`Quantile.0.975(R)`, type = "l")
```

## Gyeonggi-do data
```{r}
# library(readxl)
# library(dplyr)
# Sys.setlocale("LC_ALL", "Korean")
# dat <- read_xlsx('data/kdca.xlsx')
# dat$보도일 <- as.Date(dat$보도일, format = "%Y-%m-%d")
# dat$발병일 <- as.Date(dat$발병일, format = "%Y-%m-%d")
# dat$진단일 <- as.Date(dat$진단일, format = "%Y-%m-%d")
# dat$신고일 <- as.Date(dat$신고일, format = "%Y-%m-%d")
# 
# local <- filter(dat, 감염경로 == "국내") 
# ## Gyeonggi-do data
# gg <- local %>%
#   filter(거주시도 == "경기") %>% 
#   group_by(신고일) %>% 
#   summarize(case = n())
# names(gg) <- c("date", "case")
# saveRDS(gg, "gyeonggi.rds")
gg <- readRDS("gyeonggi.rds")
# fill in the missing dates
dat1 <- data.frame(t = seq(min(gg$date), max(gg$date), by = "day"), I = 0)
dat2 <- data.frame(t = gg$date, I = gg$case)
for (i in 1:nrow(dat1)) {
  id <- which(dat2$t == dat1$t[i])
  if(length(id) > 0){
    dat1[i, 2] <- dat2[id, 2]
  }
}

dat <- data.frame(I = round(dat1$I))

window <- 7
tstart <- seq(3, nrow(dat)- window )
tend <- tstart + window
cori_R <- 
  estimate_R(incid = dat, method = c("parametric_si"), 
             config = make_config(list(mean_si = gt_shape / gt_rate, 
                                       std_si = sqrt(gt_shape / (gt_rate^2)), 
                                       t_start = tstart, t_end= tend)))
plot(cori_R)
# plot(1:length(Rt), Rt, col = 2)
plot(tstart, cori_R$R$`Median(R)`, type = "l")
lines(tstart, cori_R$R$`Quantile.0.025(R)`, type = "l", lty = 2, col = 3)
lines(tstart, cori_R$R$`Quantile.0.975(R)`, type = "l", lty = 2, col = 3)
abline(h = 1, col = 2)
```


